---
title: "Hard Work 11"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    code_folding: hide
---

<style>

</style>

## Instructions

1. Study Sections 11.4 (first) and then 11.3 -- "Regression Trees and Robust Regression."

2. Attempt and submit at least <span id=points style="padding-left:0px;">{44}</span> Hard Work Points by Saturday at 11:59 PM.    
<span id=note>Over <span id=points style="padding-left:0px;">{47}</span> gets you {+1} Final Exam Point.</span>    


## Reading Points <span id=headpoints>{19} Possible</span>

### Section 11.4 <span id=rrecpoints>{7}</span><span id=report>{ / }</span>


### Section 11.3 <span id=rrecpoints>{12}</span><span id=report>{ / }</span>




## Theory Points <span id=headpoints>{4} Possible</span>

### Problem 11.2 <span id=points>{2}</span><span id=report>{ / }</span>
 
### Problem 11.4 <span id=points>{2}</span><span id=report>{ / }</span>

## Application Points <span id=headpoints>{25} Possible</span>


<a id=datalink style="font-size:.9em;" target="_blank" href=http://www.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/index.html>Data Files</a>

### Problem CO2 <span id=recpoints>{10}</span><span id=report>{ / }</span>

Create a graphic similar to Figure 11.9 of the textbook for the CO2 data. 
Interpret your graphic.

```{r, message=FALSE, warning=FALSE}
# Install the party library
# Run once: install.packages("party")
library(party)
```

```{r, eval=FALSE}
# Set eval=TRUE to run this code.
ctree <- ctree(uptake ~ conc + Treatment + Type, data=CO2)
plot(ctree, type="simple")

# Hint
palette(c("#FBB4AE", "#B3CDE3", "#CCEBC5", "#DECBE4"))
plot(uptake ~ conc, data=CO2, col=interaction(Type,Treatment), pch=16)
lines(c(400, 600), c(15,15), lwd=3, col="#DECBE4")
```


### Problem 11.11 <span id=recpoints>{7}</span><span id=report>{ / }</span> 

```{r}
# Code to read in the data
# It has to be merged from the 1.20 and 8.15 data files.
p11.11 <- read.table("http://www.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%2011%20Data%20Sets/CH11PR11.txt", header=FALSE)

p1.20 <- read.table("http://www.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%201%20Data%20Sets/CH01PR20.txt", header=FALSE)

p11.11 <- rbind(p1.20, p11.11)

# Give it nice column names for conevience:
colnames(p11.11) <- c("Y","X1")

# Show the data:
# p11.11
```

### Problem 11.12 <span id=recpoints>{8}</span><span id=report>{ / }</span> 

```{r}
# Code to read in the data
# It has to be merged from the 1.20 and 8.15 data files.
p11.12 <- read.table("http://www.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%2011%20Data%20Sets/CH11PR12.txt", header=FALSE)

# Give it nice column names for conevience:
colnames(p11.12) <- c("Y","X1")

# Show the data:
# p11.12
```





<footer>
</footer>



 
 <style>
 #points {
  font-size:.8em;
  padding-left:5px;
  font-weight:bold; 
  color:#317eac;
}

#recpoints {
  font-size:.8em;
  padding-left:5px;
  font-weight:bold; 
  color:#7eac31;
}

#rrecpoints {
  font-size:1em;
  padding-left:5px;
  font-weight:bold; 
  color:#7eac31;
}

#report {
  font-size:.7em;
  padding-left:15px;
  font-weight:normal; 
  color:#5a5a5a;
}

#datalink {
  font-size:.5em;
  color:#317eac;
  padding-left:5px;
}

#headnote {
  font-size:.6em;
  color:#787878;
}

#note {
  font-size:.8em;
  color:#787878;
}

#headpoints {
 font-size:12pt;
 color: #585858; 
 padding-left: 15px;
}
 </style>

 

 

 