---
title: "Personal Regression Application"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    code_folding: hide
---
```{r, include=FALSE}
library(tidyverse)
```

<style>

</style>

## Instructions

1. Use regression to help you find meaningful insights about your own life.

    * You could study your spending and earning habits.
    * You could consider your health habits as recorded by your phone or other electronic gadget.
    * You could consider the impacts of social media usage measured by number of posts, time spent, or words written, compared to your success in school each semester.
    * You could weigh your future job opportunities, graduate school, or other options against each other to find a "best choice" for you.
    * You could even take a second attempt at the Project 3 analysis, and follow up with more detail that better reflects your actual vehicle scenario. This would have to differ rather significantly from Project 3 though.

<br />

2. Your work should show breadth and depth of knowledge in the things we have learned this semester. 

## Grading

1. Breadth of Knowledge <span id=points style="padding-left:0px;">{10} Points</span>

    * What percentage of regression concepts from this semester does your analysis demonstrate?

<br />

2. Depth of Knowledge <span id=points style="padding-left:0px;">{10} Points</span>

    * Does your analysis show great expertise in at least a few of the regression topics we have covered this semester? Specifically, you need to demonstrate good ability with: the mathematical model statement, hypothesis testing, confidence or prediction intervals, and transformations.
    
## Background 

I wanted to take a look at how much electricity I use and I how much I spend relative the the temperature outside. This will help my wife and I make decesions regarding our budget and perhaps we can forecast just how much money to set aside every month. I looked at my electricty usage over the past few years to develop a simple linear model. 

## Data

Let's read in my data from Rocky Mountain Power Company:

```{r, message=FALSE}
bill <- read_csv(here::here("project_04/bill.csv")) %>% 
  rename(temp = `Average Temperature`, bill = `Bill Amount`)
```

The data came really clean so there will be little cleaning to do. Let's make a quick visualization of the temperature vs cost:

```{r}
bill %>% 
  ggplot(aes(x = temp, y = bill)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  theme_classic()
```

As we can see there are a few outliers in these data. Let's run a model and check our assumptions. 

```{r}
bill.lm <- lm(bill ~ temp, data = bill)

summary(bill.lm) %>% 
  pander::pander()

plot(bill.lm, which = c(1:2,4))
```

As we can see the assumptions are broken but not terribly so. Let's try a robust regression to smooth things out a bit. 

```{r}
bill.rlm <- MASS::rlm(bill ~ temp, data = bill)

summary(bill.rlm)
```

Okay, given a level of significance of 0.05 it appears that temperature is significant and that as temperature rises, my electricity bill also rises by about 25%. 

I guess we need to lay off the air-conditioning in the summer. 

*Sorry for the lackluster project, my original idea kind of fell apart last minute as the data I was gathering was uncleanable*
    

<footer>
</footer>



 
 <style>
 #points {
  font-size:.8em;
  padding-left:5px;
  font-weight:bold; 
  color:#317eac;
}

#recpoints {
  font-size:.8em;
  padding-left:5px;
  font-weight:bold; 
  color:#7eac31;
}

#rrecpoints {
  font-size:1em;
  padding-left:5px;
  font-weight:bold; 
  color:#7eac31;
}

#report {
  font-size:.7em;
  padding-left:15px;
  font-weight:normal; 
  color:#5a5a5a;
}

#datalink {
  font-size:.5em;
  color:#317eac;
  padding-left:5px;
}

#headnote {
  font-size:.6em;
  color:#787878;
}

#note {
  font-size:.8em;
  color:#787878;
}

#headpoints {
 font-size:12pt;
 color: #585858; 
 padding-left: 15px;
}
 </style>

 

 

 